---
title: "ADF Individual Assignment Report"
author: "Luis F. Barros - 464422"
date: "Last compiled on `r Sys.Date()`"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: 
            collapsed: false
        number_sections: true
        code_folding: hide
        theme: cerulean
editor_options: 
  chunk_output_type: console
---
# Introduction 

For this notebook, it is needed to load an R.Data file containing all the raw 
information from Twitter that I got from "collect_data.R" script. 
The output should be another .RData file with the preprocessed data that has the data converted
into data frames ready to analyze.

# Data aquisition and processing

# Data Analysis 

Load Libraries 
```{r, message=FALSE }
library("httr")
library("jsonlite")
library("tidyverse")
library("twitteR")
library("rtweet")
library("plyr")
library("dplyr")
library("data.table")
library("tidytext")
library("textdata")
library("udpipe")
library("ggplot2")
library("ggplot")
library("hrbrthemes")
library("lattice")
library("gridExtra")
library("igraph")
library("ggraph")
```

Load Objects
```{r}
# Followers tweets
  load("text.tweets.obrador.followers.RData")
  load("text.tweets.pinera.followers.RData")
  load("text.tweets.lenin.followers.RData")
  load("text.tweets.bukele.followers.RData")
  
# Presidents Tweets
  load("text.tweets.obrador.RData")
  load("text.tweets.pinera.RData")
  load("text.tweets.lenin.RData")
  load("text.tweets.bukele.RData")
  
# Presidents Timelines
  load("presidents.timelines.RData")

```




## Data visualizations

Bind all the tweets texts into a single df -- change to data processing?
```{r}
# For followers tweets
df.tweets.visualization <- rbind(text.tweets.obrador.followers, text.tweets.pinera.followers)
df.tweets.visualization <- rbind(df.tweets.visualization, text.tweets.lenin.followers)
df.tweets.visualization <- rbind(df.tweets.visualization, text.tweets.bukele.followers )
df.tweets.visualization$line <- NULL

# For presidents tweets
df.presidents.tweets.visualization <- rbind(text.tweets.obrador, text.tweets.lenin)
df.presidents.tweets.visualization <- rbind(df.presidents.tweets.visualization, 
                                            text.tweets.pinera)
df.presidents.tweets.visualization <- rbind(df.presidents.tweets.visualization,
                                            text.tweets.bukele)
df.presidents.tweets.visualization$line <- NULL
```

Make a single histogram of the tweet lenghts 
```{r}
# For followers tweets 
hist.followers.tweet.lenghts <- ggplot(df.tweets.visualization) + 
  geom_histogram(aes(x = length, fill = doc_id) , alpha = 0.6, bins = 30)
hist.followers.tweet.lenghts
```

```{r}
# For presidents tweets
hist.presidents.tweet.lenghts <- ggplot(df.presidents.tweets.visualization) + 
  geom_histogram(aes(x = length, fill = doc_id) , alpha = 0.6, bins = 30)
hist.presidents.tweet.lenghts
```


Basics statistics of the tweeter accounts of the presidents 
```{r}
average_tweet_lenght  <- mean(presidents.timelines$display_text_width)
average_tweet_lenght
```

```{r}
average_retweet_count <- mean(presidents.timelines$retweet_count)
average_retweet_count
```

Group by president and get those calculations as well then display into a plot 
```{r}
presidents.timelines <- as.data.table(presidents.timelines)
presidents.stats <-     presidents.timelines[ , list(avg_tweet_length = mean(display_text_width),
                                                     avg_retweets = mean(retweet_count)), 
                                                     by = screen_name]
presidents.stats
```

Bar Chart - Presidents average Tweet lengths 
```{r}
ggplot(presidents.stats) + geom_col(aes(x = screen_name, y = avg_tweet_length)) 
```

Bar Chart - Presidents average re-tweets
```{r}
ggplot(presidents.stats) + geom_col(aes(x = screen_name, y = avg_retweets)) 
```


## Text Mining

Given that the tweets we will analyze are in spanish language, we will use a publicly available Natural Language Processing (NLP) package for R, called "UDPipe". This package has over 65 different languages available that allow to do advances text mining analyses. 

In this section, the text frequencies of the tweets will be computed.

To start the analysis, the "Udpipie" package needs to be installed and loaded. Then we proceed to specifying our NLP model and defining the language. In this case, as the tweets we are analyzing were filtered to be in Spanish, we specify this for out NLP model.

Define the language of the analysis and download model
```{r}
udmodel <- udpipe_download_model(language = "spanish")
udmodel
```

## Analysis of followers tweets

Deploy the package in the data frames that contain the presidents followers tweets.
This will take every tweet, and create a different row for every word and character that appears in the tweet. For every word (token), additional information of that token is added. Relevant for this analysis is the "upos" column, which tells us what type of word is it, for instance, a verb, adjective, pronoun and so on.

We proceed to deploy the model into all the presidents followers tweets individually
```{r}
udmodel.obrador.followers        <- udpipe(x = text.tweets.obrador.followers$text, 
                                    object = udmodel)
udmodel.obrador.followers$doc_id <- "Mexico_followers"

udmodel.pinera.followers         <- udpipe(x = text.tweets.pinera.followers$text, 
                                    object = udmodel)
udmodel.pinera.followers$doc_id  <- "Chile_followers"

udmodel.lenin.followers          <- udpipe(x = text.tweets.lenin.followers$text, 
                                    object = udmodel)
udmodel.lenin.followers$doc_id   <- "Ecuador_followers"

udmodel.bukele.followers         <- udpipe(x = text.tweets.bukele.followers$text, 
                                    object = udmodel)
udmodel.bukele.followers$doc_id  <- "El_Salvador_followers"

```

Then the dataframes are combinedinto a single data frame to then filter more easily later on
```{r}
umodel.presidents.followers <- rbind(udmodel.obrador.followers, udmodel.pinera.followers)
umodel.presidents.followers <- rbind(umodel.presidents.followers,udmodel.lenin.followers)
umodel.presidents.followers <- rbind(umodel.presidents.followers,udmodel.bukele.followers)
head(umodel.presidents.followers)
```

As we can see, there are a lot of punctuation signs, delimiters, conjunctions and other words which are not relevant to look at. Thus, we filter to only include adjectives and nouns.
Also, by using pattern matching we can exclude all words that are mentions to other people by excluding words that have an "@" symbol.

```{r}
umodel.presidents.followers.filtered <- filter(umodel.presidents.followers,
                                               upos == "ADJ" | upos == "NOUN")
umodel.presidents.followers.filtered <- filter(umodel.presidents.followers.filtered,
                                               !grepl("@", token))
```
To check the definitions of the upos categories, check
https://universaldependencies.org/u/pos/index.html

Now that we have the relevant terms, we can proceed to calculate the text frequency


Term Frequency of followers tweets
```{r}
# Mexico
term.frequency.mexico.followers <- umodel.presidents.followers.filtered %>% 
                                      filter(doc_id == "Mexico_followers")
term.frequency.mexico.followers <- txt_freq(term.frequency.mexico.followers$token)
term.frequency.mexico.followers$country <- "Mexico"

# Chile
term.frequency.chile.followers <- umodel.presidents.followers.filtered %>% 
                                     filter(doc_id == "Chile_followers")
term.frequency.chile.followers <- txt_freq(term.frequency.chile.followers$token)
term.frequency.chile.followers$country <- "Chile"


# Ecuador
term.frequency.ecuador.followers <- umodel.presidents.followers.filtered %>% 
                                      filter(doc_id == "Ecuador_followers")
term.frequency.ecuador.followers <- txt_freq(term.frequency.ecuador.followers$token)
term.frequency.ecuador.followers$country <- "Ecuador"

# El Salvador
term.frequency.elsalvador.followers <- umodel.presidents.followers.filtered %>% 
                                         filter(doc_id == "El_Salvador_followers")
term.frequency.elsalvador.followers <- txt_freq(term.frequency.elsalvador.followers$token)
term.frequency.elsalvador.followers$country <- "El_Savador"
```

Visualize top 20 terms used by presidents followers 

```{r}
# Mexico
factor(term.frequency.mexico.followers$key, 
                    levels = rev(term.frequency.mexico.followers$key))
      barchart(key ~ freq, data = head(term.frequency.mexico.followers, 20), col = "cadetblue", 
         main = "Most occurring terms", xlab = "Freq")
```


```{r}
# Chile
factor(term.frequency.chile.followers$key, 
                    levels = rev(term.frequency.chile.followers$key))
barchart(key ~ freq, data = head(term.frequency.chile.followers, 20), col = "cadetblue", 
         main = "Most occurring terms", xlab = "Freq")
```


```{r}
# Ecuador
factor(term.frequency.ecuador.followers$key, 
                    levels = rev(term.frequency.ecuador.followers$key))
barchart(key ~ freq, data = head(term.frequency.ecuador.followers, 20), col = "cadetblue", 
         main = "Most occurring terms", xlab = "Freq")
```


```{r}
# El Salvador
factor(term.frequency.elsalvador.followers$key, 
                    levels = rev(term.frequency.elsalvador.followers$key))
barchart(key ~ freq, data = head(term.frequency.elsalvador.followers, 20), col = "cadetblue", 
         main = "Most occurring terms", xlab = "Freq")

```

Get emotions from public defined dictionaries of words and its emotions
```{r}
link <- "https://raw.githubusercontent.com/JoseCardonaFigueroa/sentiment-analysis-spanish/master/data/emotions.csv"
spanish_emotions <- read_csv(link)
# Renaming columns
names(spanish_emotions)[1] <- "key"
names(spanish_emotions)[2] <- "emotion"
```

```{r}
# Merging rows into a single data frame
followers.term.frequncies <- rbind(term.frequency.mexico.followers, term.frequency.chile.followers)
followers.term.frequncies <- rbind(followers.term.frequncies, term.frequency.ecuador.followers)
followers.term.frequncies <- rbind(followers.term.frequncies, term.frequency.elsalvador.followers)
```


Paste the emotion to the Term Frequency df
```{r}
sentiment.score.followers <- left_join(followers.term.frequncies, spanish_emotions, by = c("key"))
sentiment.score.followers <- as.data.table(na.omit(sentiment.score.followers))
sentiment.score.followers <- sentiment.score.followers[ ,list(tot_freq = sum(freq), terms_per_emotion = .N),
                                                    by = c("emotion", "country")]
```

Visualize prevalence of emotions of followers
```{r}
# Stacked
sentiment_followers_stacked_plot <- ggplot(sentiment.score.followers, 
                                aes(fill=country, y=tot_freq, x=emotion)) + 
                                geom_bar(position="stack", stat="identity") +
                                  ggtitle("Sentiment of Followers")

sentiment_followers_stacked_plot
```


## Analysis of presidents tweets
```{r}
udmodel.obrador        <- udpipe(x = text.tweets.obrador$text, 
                                    object = udmodel)
udmodel.obrador$doc_id <- "Mexico"

udmodel.pinera         <- udpipe(x = text.tweets.pinera$text, 
                                    object = udmodel)
udmodel.pinera$doc_id  <- "Chile"

udmodel.lenin          <- udpipe(x = text.tweets.lenin$text, 
                                    object = udmodel)
udmodel.lenin$doc_id   <- "Ecuador"

udmodel.bukele         <- udpipe(x = text.tweets.bukele$text, 
                                    object = udmodel)
udmodel.bukele$doc_id  <- "El Salvador"

```

Merge them into a single data frame to then filter more easily
```{r}
umodel.presidents <- rbind(udmodel.obrador, udmodel.pinera)
umodel.presidents <- rbind(umodel.presidents,udmodel.lenin)
umodel.presidents <- rbind(umodel.presidents,udmodel.bukele)
```

```{r}
umodel.presidents.filtered <- filter(umodel.presidents, 
                                             upos == "ADJ" | upos == "NOUN" |
                                             upos == "PROPN" | upos == "INTJ")
umodel.presidents.filtered <- filter(umodel.presidents.filtered, 
                                             !grepl("@", token))
```

Term Frequency of president Tweets
```{r}
# Mexico
term.frequency.mexico.president            <- umodel.presidents.filtered %>% 
                                              filter(doc_id == "Mexico")
term.frequency.mexico.president            <- txt_freq(term.frequency.mexico.president$token)
term.frequency.mexico.president$president  <- "Mexico"

# Chile
term.frequency.chile.president             <- umodel.presidents.filtered %>% 
                                              filter(doc_id == "Chile")
term.frequency.chile.president             <- txt_freq(term.frequency.chile.president$token)
term.frequency.chile.president$president   <- "Chile"

# Ecuador 
term.frequency.ecuador.president           <- umodel.presidents.filtered %>% 
                                              filter(doc_id == "Ecuador")
term.frequency.ecuador.president           <- txt_freq(term.frequency.ecuador.president$token)
term.frequency.ecuador.president$president <- "Ecuador"

# El Salvador
term.frequency.elsalvador.president         <- umodel.presidents.filtered %>% 
                                              filter(doc_id == "El Salvador")
term.frequency.elsalvador.president         <- txt_freq(term.frequency.elsalvador.president$token)
term.frequency.elsalvador.president$president <- "El Salvador"
```

```{r}
# Merging presidents frequnecies 
presidents.term.frequencies <- rbind(term.frequency.mexico.president, term.frequency.chile.president)
presidents.term.frequencies <- rbind(presidents.term.frequencies,term.frequency.ecuador.president)
presidents.term.frequencies <- rbind(presidents.term.frequencies,term.frequency.elsalvador.president)
```

```{r}
# Joing the emtion to the words
sentiment.score.presidents <- left_join(presidents.term.frequencies, spanish_emotions, by = "key")
sentiment.score.presidents <-  as.data.table(na.omit(sentiment.score.presidents))
sentiment.score.presidents <- sentiment.score.presidents[ ,list(tot_freq = sum(freq), 
                                                      terms_per_sentiment = .N), 
                                                      by = c("emotion", "president")]
```

```{r}
# Stacked bar chart
sentiment_presidents_stacked_plot <- ggplot(sentiment.score.presidents, 
                                aes(fill=president, y=tot_freq, x=emotion)) + 
                                geom_bar(position="stack", stat="identity") + 
                                ggtitle("Sentiment of Presidents")

sentiment_presidents_stacked_plot
```

Compare the plots
```{r}
president.and.followers.plot <- grid.arrange(sentiment_presidents_stacked_plot, sentiment_followers_stacked_plot, nrow=2)
president.and.followers.plot
```

## Co-occurrence Analysis
```{r}
co_occurence_presidents <- cooccurrence(umodel.presidents.filtered, term = "token", group = "token_id")

co_occurence_followers <- cooccurrence(umodel.presidents.followers.filtered,
                                       term = "token", group = "token_id")
```

Visualize Network of co-occurence 
```{r}
wordnetwork_followers <- head(co_occurence_followers, 30)
wordnetwork_followers <- graph_from_data_frame(wordnetwork_followers)
ggraph(wordnetwork_followers, layout = "fr") +
  geom_edge_link(aes(width = 30, edge_alpha = 30), edge_colour = "pink") +
  geom_node_text(aes(label = name), col = "darkgreen", size = 4) +
  theme_graph(base_family = "Arial Narrow") +
  theme(legend.position = "none") +
  labs(title = "Cooccurrences within sentence", subtitle = "Followers Tweets")
```



# Discussion 

## Implications 
Most frequent terms between presidents and followers

Important differences 

Insights based on sentiment analysis

## Limitations 
- Small sample size due to procesing time
- Sample is biased in the last 9 days of when tweets were collected, so not representative
  - empahsis in covid, vaccination, lockdown,

# References 

https://www.tidytextmining.com/sentiment.html
https://bnosac.github.io/udpipe/docs/doc5.html
https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html
http://www.consulta.mx/index.php/encuestas-e-investigaciones/el-mundo
https://cran.r-project.org/web/packages/udpipe/vignettes/udpipe-annotation.html#udpipe_the_R_package
